### 2020-10-15 priority-cp.tex
Compressed PCP metatheory.

\subsubsection*{Subject reduction}
Reduction in our version of PCP preserve types. Since our reduction relation is a strict subset of the reduction relation in the original~\cite{dardhagay18}, we defer to their proofs:
\begin{compacttheorems}
  \begin{restatablelemma}{lempcpsubjectcongruence}[Subject Congruence, $\pcp{\equiv}$]
    \label{lem:pcp-subject-congruence}
    If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\equiv\tm{Q}}$,
    then $\pcp{\seq{Q}{\ty{\Gamma}}}$.
  \end{restatablelemma}
  \begin{restatabletheorem}{thmpcpsubjectreduction}[Subject Reduction, $\pcp{\red}$]
    \label{thm:pcp-subject-reduction}
    If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\red\tm{Q}}$, then $\pcp{\seq{Q}{\ty{\Gamma}}}$.
  \end{restatabletheorem}
\end{compacttheorems}

\subsubsection*{Progress}
Reduction in our version of PCP satisfies progress:
\begin{restatabletheorem}{thmpcpclosedprogress}[Progress, $\pcp{\red}$]
  \label{thm:pcp-closed-progress}
  \hfill\\%newline before theorem statement
  If $\pcp{\seq{P}{\emptyenv}}$, then either $\pcp{\tm{P}=\tm{\halt}}$ or there exists a $\pcp{\tm{Q}}$ such that $\pcp{\tm{P}\red\tm{Q}}$.
\end{restatabletheorem}
\begin{proof}
    Our proof follows the reasoning by Kobayashi~\cite{kobayashi06} used in the proof of deadlock freedom for closed processes (Theorem 2). The details can be found in~\Cref{prf:thm-pcp-closed-progress}.
\end{proof}

\subsubsection*{Translation}
The translation on types from PCP to PGV, $\ty{\cpgvT{\cdot}}$, is:
\[
  \begin{array}{lcl}
    \ty{\cpgvT{\pcp{\tytens[\cs{o}]{A}{B}}}}
    & = & \pgv{\ty{\tysend[\cs{o}]{\co{\cpgvT{A}}}{\cpgvT{B}}}}
    \\
    \ty{\cpgvT{\pcp{\typlus[\cs{o}]{A}{B}}}}
    & = & \pgv{\ty{\tyselect[\cs{o}]{\cpgvT{A}}{\cpgvT{B}}}}
  \end{array}
  \;
  \begin{array}{lcl}
    \ty{\cpgvT{\pcp{\tyone[\cs{o}]}}}
    & = & \ty{\pgv{\tyends[\cs{o}]}}
    \\
    \ty{\cpgvT{\pcp{\tynil[\cs{o}]}}}
    & = & \pgv{\ty{\tyselectemp[\cs{o}]}}
  \end{array}
  \;
  \begin{array}{lcl}
    \ty{\cpgvT{\pcp{\typarr[\cs{o}]{A}{B}}}}
    & = & \pgv{\ty{\tyrecv[\cs{o}]{\cpgvT{A}}{\cpgvT{B}}}}
    \\
    \ty{\cpgvT{\pcp{\tywith[\cs{o}]{A}{B}}}}
    & = & \pgv{\ty{\tyoffer[\cs{o}]{\cpgvT{A}}{\cpgvT{B}}}}
  \end{array}
  \;
  \begin{array}{lcl}
    \ty{\cpgvT{\pcp{\tybot[\cs{o}]}}}
    & = & \ty{\pgv{\tyendr[\cs{o}]}}
    \\
    \ty{\cpgvT{\pcp{\tytop[\cs{o}]}}}
    & = & \pgv{\ty{\tyofferemp[\cs{o}]}}
  \end{array}
\]
There are two separate translations on processes. The first, $\tm{\cpgvC{\cdot}}$, translates processes to \emph{configurations}. It translates as much as possible of the process to the corresponding configuration constructs. Once it hits the first action, it switches to the second translation, $\tm{\cpgvM{\cdot}}$, which translates processes to \emph{terms}:
\[
  \begin{array}{lcl}
    \pcp{\tm{\cpgvC{\res{x}{y}{P}}}}
    & = & \pgv{\tm{\res{x}{y}{\cpgvC{P}}}} \\
    \pcp{\tm{\cpgvC{\ppar{P}{Q}}}}
    & = & \pgv{\tm{\ppar{\cpgvC{P}}{\cpgvC{Q}}}} \\
    \pcp{\tm{\cpgvC{\send{x}{y}{P}}}}
    & = & \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\send\;\pair{z}{x}}{\cpgvM{P}})}}} \\
    \pcp{\tm{\cpgvC{\inl{x}{P}}}}
    & = & \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inl{y}}{x})}{z}}{\cpgvM{P}})}}} \\
    \pcp{\tm{\cpgvC{\inr{x}{P}}}}
    & = & \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inr{y}}{x})}{z}}{\cpgvM{P}})}}} \\
    \pcp{\tm{\cpgvC{P}}}
    & = & \pgv{\tm{\child{\cpgvM{P}}}},\hfill\text{if none of the above apply}\\
    \\
    \pcp{\tm{\cpgvM{\link{x}{y}}}}
    & = & \pgv{\tm{\link\;{\pair{x}{y}}}} \\
    \pcp{\tm{\cpgvM{\res{x}{y}{P}}}}
    & = & \pgv{\tm{\letpair{x}{y}{\new}{\cpgvM{P}}}} \\
    \pcp{\tm{\cpgvM{\ppar{P}{Q}}}}
    & = & \pgv{\tm{\andthen{\spawn\;{(\lambda\unit.\cpgvM{P})}}{\cpgvM{Q}}}} \\
    \pcp{\tm{\cpgvM{\halt}}}
    & = & \pgv{\tm{\unit}} \\
    \pcp{\tm{\cpgvM{\close{x}{P}}}}
    & = & \pgv{\tm{\andthen{\close\;{x}}{\cpgvM{P}}}} \\
    \pcp{\tm{\cpgvM{\wait{x}{P}}}}
    & = & \pgv{\tm{\andthen{\wait\;{x}}{\cpgvM{P}}}} \\
    \pcp{\tm{\cpgvM{\send{x}{y}{P}}}}
    & = & \pgv{\tm{\letpair{y}{z}{\new}{\letbind{x}{\send\;{\pair{z}{x}}}{\cpgvM{P}}}}} \\
    \pcp{\tm{\cpgvM{\recv{x}{y}{P}}}}
    & = & \pgv{\tm{\letpair{y}{x}{\recv\;{x}}{\cpgvM{P}}}} \\
    \pcp{\tm{\cpgvM{\inl{x}{P}}}}
    & = & \pgv{\tm{\letbind{x}{\select{\labinl}\;{x}}{\cpgvM{P}}}} \\
    \pcp{\tm{\cpgvM{\inr{x}{P}}}}
    & = & \pgv{\tm{\letbind{x}{\select{\labinr}\;{x}}{\cpgvM{P}}}} \\
    \multicolumn{3}{l}{\pcp{\tm{\cpgvM{\offer{x}{P}{Q}}}}} \\
    & = & \pgv{\tm{\offer{x}{x}{\cpgvM{P}}{x}{\cpgvM{Q}}}} \\
    \pcp{\tm{\cpgvM{\absurd{x}}}}
    & = & \pgv{\tm{\offeremp{x}}}
  \end{array}
\]
The third, fourth, and fifth cases for $\tm{\cpgvC{\cdot}}$ translate the $\nu$-binders implicit in the encoding of bound sending to configuration constructs, which is required for completeness of the operational correspondence.
Typing environments are translated pointwise, and sequents $\pcp{\seq{P}{\ty{\Gamma}}}$ are translated as $\pgv{\cseq[\child]{\ty{\cpgvT{\ty{\Gamma}}}}{\cpgvC{P}}}$.

@@

\subsubsection*{Type preservation}
The translations $\tm{\cpgvC{\cdot}}$ and $\tm{\cpgvM{\cdot}}$ are preserve typing. The proofs are standard inductions. The details can be found in~\cref{prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-terms-preservation}:
\begin{compacttheorems}
  \begin{restatablelemma}{lempcptopgvtermspreservation}[Preservation, ${\tm{\cpgvM{\cdot}}}$]
    \label{lem:pcp-to-pgv-terms-preservation}
    If $\pcp{\seq{P}{\ty{\Gamma}}}$, then $\pgv{\tseq[\cs{p}]{\ty{\cpgvT{\Gamma}}}{\cpgvM{P}}{\tyunit}}$.
  \end{restatablelemma}
  \begin{restatabletheorem}{thmpcptopgvconfspreservation}[Preservation, ${\tm{\cpgvC{\cdot}}}$]
    \label{thm:pcp-to-pgv-confs-preservation}
    If $\pcp{\seq{P}{\ty{\Gamma}}}$, then $\pgv{\cseq[\child]{\ty{\cpgvT{\Gamma}}}{\cpgvC{P}}}$.
  \end{restatabletheorem}
\end{compacttheorems}

\subsubsection*{Operational correspondence}
Finally, we show that PCP and PGV are in operational correspondence along our translation. First, we link our two translations, and show that $\tm{\cpgvM{P}}$ \emph{must} reduce to $\tm{\cpgvC{P}}$, \ie $\tm{\cpgvC{\cdot}}$ is equivalent to translating with $\tm{\cpgvM{\cdot}}$ and performing several reduction steps:
\begin{compacttheorems}
  \begin{restatablelemma}{lempcptopgvcpgvMtocpgvC}
    \label{lem:pcp-to-pgv-cpgvM-to-cpgvC}
    For any $\pcp{\tm{P}}$, either:
    \begin{itemize}
    \item $\pgv{\tm{\child\;\cpgvM{P}}=\tm{\cpgvC{P}}}$; or
    \item   $\pgv{\tm{\child\;\cpgvM{P}}\cred^+\tm{\cpgvC{P}}}$, and for any $\pgv{\tm{\conf{C}}}$, if $\pgv{\tm{\child\;\cpgvM{P}}\cred\tm{\conf{C}}}$, then $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{P}}}$.
    \end{itemize}
  \end{restatablelemma}
  \begin{proof}
    By induction on the structure of $\pcp{\tm{P}}$. Details in \cref{prf:lem-pcp-to-pgv-cpgvM-to-cpgvC}.
  \end{proof}
  \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencecompleteness}%
    [Operational Correspondence, Completeness, ${\tm{\cpgvC{\cdot}}}$]
    \label{thm:pcp-to-pgv-operational-correspondence-completeness}
    \hfill\\%newline before theorem statement
    If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\red\tm{Q}}$,
    then $\pgv{\tm{\cpgvC{P}}\cred^+\tm{\cpgvC{Q}}}$.
  \end{restatabletheorem}
  \begin{proof}
    By induction on the derivation of $\pcp{\tm{P}\red\tm{Q}}$. We apply \cref{lem:pcp-to-pgv-cpgvM-to-cpgvC} for the cases in which the translation switches from $\tm{\cpgvC{\cdot}}$ to $\tm{\cpgvM{\cdot}}$, \ie under actions. We apply \cref{lem:pcp-to-pgv-confs-operational-correspondence-equiv} for \LabTirName{E-LiftSC}. Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-completeness}.
  \end{proof}
  \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencesoundness}%
    [Operational Correspondence, Soundness, ${\tm{\cpgvC{\cdot}}}$]
    \label{thm:pcp-to-pgv-operational-correspondence-soundness}
    \hfill\\%newline before theorem statement
    If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$, there exists a $\tm{Q}$ such that $\pcp{\tm{P}\red^+\tm{Q}}$ and $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{Q}}}$
  \end{restatabletheorem}
  \begin{proof}
    By induction on the derivation of $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$.\simon{Surprising that you can get it by the derivation of the reduction relation---normally, 'soundness' would be proved by shunting into a canonical form and then proceeding by induction the structure of the term / induction on typing derivations.}
    Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-soundness}.
  \end{proof}
\end{compacttheorems}

>>>>

\subsubsection*{Metatheory}
Reduction in our version of PCP preserve types. Since our reduction relation is a strict subset of the reduction relation in the original~\cite{dardhagay18}, we defer to their proofs. We prove progress for our version of PCP, see~\cref{prf:thm-pcp-closed-progress}.

\subsubsection*{Translation}
We illustrate the relation between PCP and PGV by giving a translation from PCP to PGV. The translation on types is defined as follows:
\[
  \begin{array}{lcl}
    \ty{\cpgvT{\pcp{\tytens[\cs{o}]{A}{B}}}}
    & = & \pgv{\ty{\tysend[\cs{o}]{\co{\cpgvT{A}}}{\cpgvT{B}}}}
    \\
    \ty{\cpgvT{\pcp{\typlus[\cs{o}]{A}{B}}}}
    & = & \pgv{\ty{\tyselect[\cs{o}]{\cpgvT{A}}{\cpgvT{B}}}}
  \end{array}
  \;
  \begin{array}{lcl}
    \ty{\cpgvT{\pcp{\tyone[\cs{o}]}}}
    & = & \ty{\pgv{\tyends[\cs{o}]}}
    \\
    \ty{\cpgvT{\pcp{\tynil[\cs{o}]}}}
    & = & \pgv{\ty{\tyselectemp[\cs{o}]}}
  \end{array}
  \;
  \begin{array}{lcl}
    \ty{\cpgvT{\pcp{\typarr[\cs{o}]{A}{B}}}}
    & = & \pgv{\ty{\tyrecv[\cs{o}]{\cpgvT{A}}{\cpgvT{B}}}}
    \\
    \ty{\cpgvT{\pcp{\tywith[\cs{o}]{A}{B}}}}
    & = & \pgv{\ty{\tyoffer[\cs{o}]{\cpgvT{A}}{\cpgvT{B}}}}
  \end{array}
  \;
  \begin{array}{lcl}
    \ty{\cpgvT{\pcp{\tybot[\cs{o}]}}}
    & = & \ty{\pgv{\tyendr[\cs{o}]}}
    \\
    \ty{\cpgvT{\pcp{\tytop[\cs{o}]}}}
    & = & \pgv{\ty{\tyofferemp[\cs{o}]}}
  \end{array}
\]
There are two separate translations on processes. The first, $\tm{\cpgvC{\cdot}}$, translates processes to \emph{configurations}. It translates as much as possible of the process to the corresponding configuration constructs. Once it hits the first action, it switches to the second translation, $\tm{\cpgvM{\cdot}}$, which translates processes to \emph{terms}:
\[
  \begin{array}{lcl}
    \pcp{\tm{\cpgvC{\res{x}{y}{P}}}}
    & = & \pgv{\tm{\res{x}{y}{\cpgvC{P}}}} \\
    \pcp{\tm{\cpgvC{\ppar{P}{Q}}}}
    & = & \pgv{\tm{\ppar{\cpgvC{P}}{\cpgvC{Q}}}} \\
    \pcp{\tm{\cpgvC{\send{x}{y}{P}}}}
    & = & \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\send\;\pair{z}{x}}{\cpgvM{P}})}}} \\
    \pcp{\tm{\cpgvC{\inl{x}{P}}}}
    & = & \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inl{y}}{x})}{z}}{\cpgvM{P}})}}} \\
    \pcp{\tm{\cpgvC{\inr{x}{P}}}}
    & = & \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inr{y}}{x})}{z}}{\cpgvM{P}})}}} \\
    \pcp{\tm{\cpgvC{P}}}
    & = & \pgv{\tm{\child{\cpgvM{P}}}},\hfill\text{if none of the above apply}\\
    \\
    \pcp{\tm{\cpgvM{\link{x}{y}}}}
    & = & \pgv{\tm{\link\;{\pair{x}{y}}}} \\
    \pcp{\tm{\cpgvM{\res{x}{y}{P}}}}
    & = & \pgv{\tm{\letpair{x}{y}{\new}{\cpgvM{P}}}} \\
    \pcp{\tm{\cpgvM{\ppar{P}{Q}}}}
    & = & \pgv{\tm{\andthen{\spawn\;{(\lambda\unit.\cpgvM{P})}}{\cpgvM{Q}}}} \\
    \pcp{\tm{\cpgvM{\halt}}}
    & = & \pgv{\tm{\unit}} \\
    \pcp{\tm{\cpgvM{\close{x}{P}}}}
    & = & \pgv{\tm{\andthen{\close\;{x}}{\cpgvM{P}}}} \\
    \pcp{\tm{\cpgvM{\wait{x}{P}}}}
    & = & \pgv{\tm{\andthen{\wait\;{x}}{\cpgvM{P}}}} \\
    \pcp{\tm{\cpgvM{\send{x}{y}{P}}}}
    & = & \pgv{\tm{\letpair{y}{z}{\new}{\letbind{x}{\send\;{\pair{z}{x}}}{\cpgvM{P}}}}} \\
    \pcp{\tm{\cpgvM{\recv{x}{y}{P}}}}
    & = & \pgv{\tm{\letpair{y}{x}{\recv\;{x}}{\cpgvM{P}}}} \\
    \pcp{\tm{\cpgvM{\inl{x}{P}}}}
    & = & \pgv{\tm{\letbind{x}{\select{\labinl}\;{x}}{\cpgvM{P}}}} \\
    \pcp{\tm{\cpgvM{\inr{x}{P}}}}
    & = & \pgv{\tm{\letbind{x}{\select{\labinr}\;{x}}{\cpgvM{P}}}} \\
    \multicolumn{3}{l}{\pcp{\tm{\cpgvM{\offer{x}{P}{Q}}}}} \\
    & = & \pgv{\tm{\offer{x}{x}{\cpgvM{P}}{x}{\cpgvM{Q}}}} \\
    \pcp{\tm{\cpgvM{\absurd{x}}}}
    & = & \pgv{\tm{\offeremp{x}}}
  \end{array}
\]
The third, fourth, and fifth cases for $\tm{\cpgvC{\cdot}}$ translate the $\nu$-binders implicit in the encoding of bound sending to configuration constructs, which is required for completeness of the operational correspondence.
Typing environments are translated pointwise, and sequents $\pcp{\seq{P}{\ty{\Gamma}}}$ are translated as $\pgv{\cseq[\child]{\ty{\cpgvT{\ty{\Gamma}}}}{\cpgvC{P}}}$.

@@

The translations $\tm{\cpgvC{\cdot}}$ and $\tm{\cpgvM{\cdot}}$ are preserve typing. The proofs are standard inductions. The details can be found in~\cref{prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-terms-preservation}:
\begin{compacttheorems}
  \begin{restatablelemma}{lempcptopgvtermspreservation}[Preservation, ${\tm{\cpgvM{\cdot}}}$]
    \label{lem:pcp-to-pgv-terms-preservation}
    If $\pcp{\seq{P}{\ty{\Gamma}}}$, then $\pgv{\tseq[\cs{p}]{\ty{\cpgvT{\Gamma}}}{\cpgvM{P}}{\tyunit}}$.
  \end{restatablelemma}
  \begin{restatabletheorem}{thmpcptopgvconfspreservation}[Preservation, ${\tm{\cpgvC{\cdot}}}$]
    \label{thm:pcp-to-pgv-confs-preservation}
    If $\pcp{\seq{P}{\ty{\Gamma}}}$, then $\pgv{\cseq[\child]{\ty{\cpgvT{\Gamma}}}{\cpgvC{P}}}$.
  \end{restatabletheorem}
\end{compacttheorems}

Our translation satisfies operational correspondence. We link our translations, and show that $\tm{\cpgvM{P}}$ \emph{must} reduce to $\tm{\cpgvC{P}}$, \ie $\tm{\cpgvC{\cdot}}$ is equivalent to translating with $\tm{\cpgvM{\cdot}}$ and performing several reduction steps. We use this fact to prove operational correspondence. The details are in~\cref{prf:lem-pcp-to-pgv-cpgvM-to-cpgvC,prf:thm-pcp-to-pgv-operational-correspondence-completeness,prf:thm-pcp-to-pgv-operational-correspondence-soundness}:
\begin{compacttheorems}
  \begin{restatablelemma}{lempcptopgvcpgvMtocpgvC}
    \label{lem:pcp-to-pgv-cpgvM-to-cpgvC}
    For any $\pcp{\tm{P}}$, either:
    \begin{itemize}
    \item $\pgv{\tm{\child\;\cpgvM{P}}=\tm{\cpgvC{P}}}$; or
    \item   $\pgv{\tm{\child\;\cpgvM{P}}\cred^+\tm{\cpgvC{P}}}$, and for any $\pgv{\tm{\conf{C}}}$, if $\pgv{\tm{\child\;\cpgvM{P}}\cred\tm{\conf{C}}}$, then $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{P}}}$.
    \end{itemize}
  \end{restatablelemma}
  \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencecompleteness}%
    [Operational Correspondence, Completeness, ${\tm{\cpgvC{\cdot}}}$]
    \label{thm:pcp-to-pgv-operational-correspondence-completeness}
    \hfill\\%newline before theorem statement
    If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\red\tm{Q}}$,
    then $\pgv{\tm{\cpgvC{P}}\cred^+\tm{\cpgvC{Q}}}$. Details in~\cref{prf:lem-pcp-to-pgv-cpgvM-to-cpgvC}.
  \end{restatabletheorem}
  % \begin{proof}
  %   By induction on the derivation of $\pcp{\tm{P}\red\tm{Q}}$. We apply \cref{lem:pcp-to-pgv-cpgvM-to-cpgvC} for the cases in which the translation switches from $\tm{\cpgvC{\cdot}}$ to $\tm{\cpgvM{\cdot}}$, \ie under actions. We apply \cref{lem:pcp-to-pgv-confs-operational-correspondence-equiv} for \LabTirName{E-LiftSC}. Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-completeness}.
  % \end{proof}
  \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencesoundness}%
    [Operational Correspondence, Soundness, ${\tm{\cpgvC{\cdot}}}$]
    \label{thm:pcp-to-pgv-operational-correspondence-soundness}
    \hfill\\%newline before theorem statement
    If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$, there exists a $\tm{Q}$ such that $\pcp{\tm{P}\red^+\tm{Q}}$ and $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{Q}}}$
  \end{restatabletheorem}
  % \begin{proof}
  %   By induction on the derivation of $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$.\simon{Surprising that you can get it by the derivation of the reduction relation---normally, 'soundness' would be proved by shunting into a canonical form and then proceeding by induction the structure of the term / induction on typing derivations.}
  %   Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-soundness}.
  % \end{proof}
\end{compacttheorems}
